Digital Privacy and Cryptography

### 2014-12-10 LaTex/paper.tex
As good as my writing abilities will take me

\begin{abstract}
Do people care about digital privacy and how can cryptography keep digital information private
\end{abstract}

@@

The development of databases, the growth of mobile industry  and the increase usage of the Internet have resulted in a huge collection of personal data. More companies are moving their Information Techonology to the cloud and are creating Data Centers which store an inconceivable amounts of data. Facebook has claimed that they have over a 100 petabytes of photos and video (Wallbank, 2012, para. 8). This along with a strong developer community has lead to anyone being able to create a web application. There are tons of web based hosting companies such as Heroku which allows any to publish their web application easily. New technologies are created yearly for all kinds of web development. There are over 250,000 web related repositories on Github alone. There are numerous plugins that enable your web application to be integrated with numerous social media websites such as Twitter and Facebook. The Internet has provided connections across the world but these connections have brought along some issues.

@@

People are constantly giving out personal information such as email, address, and credit card daily through various websites. All kinds of information is being tracked when browsing the Internet through Google analytics. With all this information being shared and stored everyday, are people concerned with its security and their privacy?

@@

Business are driven by profit and will look for ways to maximize profit in any manner, if there is not government intervention. Sadly there is a lack of legislation dealing with the Internet. This is largely due to the how fast web techonologies have grown and how slow the political systems moves. Most of contract law is obsolete when it comes to the Internet and Internet transactions. One example is Canada's Anti-Spam Legislation (CASL) which was passed in Decemeber 2010 but not enforced until July 1, 2014(Government of Canada, 2014, para. 1). It took about a decade and a half to create a spam law and another four years to implement it. The lack of legislation allows for companies to freely deal with their data, and user agreements with little government interference. The goverment is doing little to ensure digital privacy and in certain cases is exploiting personal information in the name of security. A recent example is the Federal Buraeu of Investigation's complaint about Apple's Iphone encryption of data where the user's password was used to encrypt the data (Schneier, 2014, para. 3-4). Encryption is more of a hassle and back doors make their job much easier.

All business have the incentive to improve profit and without external incentives will strictly focus on profits. The two greatest external forces on corporations are the government and consumer purchasing power. The government's pressure on companies is non-existence and what force that is present is outdated. The public is less concerned about with drawbacks of the data privacy and more concerned with the benefits of the application. Most people ask what can it do for me instead of how does it work and what security measures are in place? The lack of incentives have lead to businesses tending to focus more what profit can be produced from personal data and not the long term impact that it could have. There is a lack of research done by the companies producing these techonologies on the users. The negative side effects are not taken into consideration during the design process. Security is usually an after thought and privacy is never considered. This is evident in a recent study done by Cisco Canada in which they found ``40 per cent of about 500 firms surveyed had security strategies'' (Blackwell, 2014, para. 2). Just last week, Sony was hacked where a significant amount of employee information was leaked. Their are estimates of ``100 terabytes of data'' leaked containing information such as ``list of employee salaries and bonuses; Social Security numbers and birth dates; HR employee performance reviews, criminal background checks and termination records'' (Zetter, 2014, para 15-16).

@@

The culture change started by the Internet has brought new social issues. Now a citizen is evaluated by their on-line social interaction. Companies will use Facebook, LinkedIn and Google to screen out potential candidates. Employers need to monitor the social lives closely because of this. One cyber attack can greatly damage one's reputation and may threaten their employment or even their employability. The other thing is the Internet does not forget about one's past. A good example of this is the recent firing of a Quebec teacher: ``the availability on the Internet of erotic films in which she acted created an entirely new context that was not ideal for our students'' (Peritz, 2014, para. 8) was the reason for her dismissal. A teacher's work that occurred over forty years ago should not be revelant to the evaluation of their teaching. This would not have been an issue if the students had just heard about it since it would just be a rumour. The actual video itself which is shared on the Internet is what the school claims to be the issue. The teacher has no control over the distribution of the content now. This is punishing someone for something that happen a long time ago in which they have no control over it. This problem will only get larger as more information is stored and moved into an age where all data becomes digitalized. It is a frightening fact that all mistakes in one's lifetime will not be forgotten and could impact someone's reputation world wide even forty years later.

The new buzz word in the business world is big data and processing big data for applications. The applications are used by business to make decision and to help better market their product. There is a big push for all companies to collect data and analyzing this data to get ahead. The collection of personal data is usually supported by the possibilities that big data analytics could provide. Their are large claims by people of all the potential benefits including big data makes people happy (Banayan, 2014, para. 4). The collection of this data has little regulations and with an increasing financial incentive it can or may cause some business to collect this information in unethical ways. The other problem is that there is a over confidence in this new field. Micheal Jordan a professor at the University of California stated that ``When you have large amounts of data, your appetite for hypotheses tends to get even larger. And if it’s growing faster than the statistical strength of the data, then many of your inferences are likely to be false. They are likely to be white noise.'' (Gomes, 2014, para. 30) . Google Flu Trends is a system Google developed for tracking the flu. ``Google Flu Trends is often held up as an exemplary use of big data''. (Lazer, 2014, pg. 1). It was reported that Google Flu Trend's prediction were twice as much as the proportion of actual doctor vists for influenze to the Center for Disease Control and Prevention (Lazer, 2014, pg. 1). This is a good case example of how big data does not have all the answers and that is a field that needs improvement and reasearch. Over confidence in it could lead to costly mislead decisions based on ``big data hubris'' (Lazer, 2014, pg.1).

@@

\section{Ways to  Change}\label{sec:developers}
It is clear that security and digital privacy need to have more focus and be strengthened. A developer needs to recognize this during the design of their system. They need to build into their system in ways that users can stay annonymous or delete private information when requested. Digital privacy should be the standard when creating web applications and not a feature.

@@

The steps of the system start by the user creating a note number called $n$ which is a random number. The users then creates a second random number $r$. The note number $n$ is then blinded by combining $n$ with $r$. This is done mathematically by $n * r^b$ where $b$ is the bank's public power which equates to one dollar. The blinded note is then transferred to the bank. The banks signs the blinded note by using its private power $\overline{b}$. The result is returned and the signed note number is obtained by dividing out the $r$ to obtain $n^{\overline{b}}$. The note can be verified that is a signed special number (Chaum, 1985, pg. 7-8).

Banks needs to keep track of transactions to ensure that no note has been used twice. This is done by having a list of all accepted transactions. A transaction is only allowed if it is not part of the list and has the appropriate signature. The underlying digital signature techniques make it difficult for anyone to forge the bank's signature $\overline{b}$. If $r$ is not disclosed then the transcation deposit cannont be linked to the withdrawel (Chaum, 1985, pg. 8-9).

@@

These anonymous remailer aim to allow users to communicate through e-mail without losing privacy. According to Ian Goldberg, David Wagner, and Eric Brewer there are four differ Versions of Type X Remailers, Type O, Type I, Type II and Type III. Each type provides addtional functionality than the previous versions. All the different versions work differently where some are anonymity systems and others are pseudonym systems.

@@

The basic idea of Onion Routing is to create an indirect path between nodes starting at the user's node and ending at the recipents. Each node in the path does not know the recipent but rather just the next node in the path. Each node does not know the path due to the way the packet is encrypted. Each node removes a level of encryption and determines which node to pass to. This path may be used so the recipent can respond which allows for pseudonymous communications. (Reed, Syverson, \& Goldschag, 1998, pg. 2)

@@

The protocol first exchanges keys using a key exchange. The data is then encrypted using a specified encyrption function such as AES. A message authencating code (MAC) is also created. The encyrpted text and the MAC is sent. The encyrption does not allow for eavesdropper to read the intended message. The MAC does not allow the user to intercept the message and change it (Sidhpurwala, 2013, para.3).

@@

 Transport Layer Security is widely used. Transport Layer Security provides an example of when standards are created then people will follow them. It is not difficult to enable a web application to support Transport Layer Security through various libraries such as OpenSSL. However, it also is an example of how security budgets are inadequate when compared to their level of importance. This was seen in the HeartBleed bug which occurred in the OpenSSL library this year. OpenSSl is used widely but developed and maintained by a small community of people. It has been stated that ``66 percent of all web servers'' run OpenSSL but OpenSSL Software Foundation receives only ``\$2,000 a year in donations'' (Warren, 2014, para. 28-30). If the support of OpenSSL was comparable to the user of the software then the bug would have been much less likely to occur.

\section{Moving Forward}\label{sec:moving-forward}
There are many more privacy enhancing technologies besides the four discusses previously. There is a vast collection of libraries that developers can leverage to help them make their applications more secure. In the early steps phases of design developers need to focus on security and how their applications deal with privacy issues. A developer has a responsibility to their user's ensure their information is secure and that the user has control over thier private information.

@@

Businesses need to create strategies to handle security and make clear in their user agreements how they store and use information collected. Facebook recently moved towards this by shortening their user agreement and making it more readable for users (Albergotti, 2014, para. 4). This was a step in the write direction but still mores need to be done. Business needs to increases security budgets which are too small. Some companies are seeing this need which can be seen  by the IAPP study which reported ``38 percent of respondents said they would likely increase their privacy budget in the next year'' (Solove, 2014, para. 2). There are some companies who have focused on creating secure systems.

\section{Businesses focusing on Security}\label{sec:real-life}
Wickr is a new messenger which is available on Google Play store, Apple App Store, and even for the desktop. It is uses peer-to-peer encryption and enables users to send encrypted messages.  According to its website it uses hashed device information with SHA256 and encrypts all data with AES256. The Wickr server only handles encrypted messages and delete the message upon delivery. The server also never directly interacts with devices identifiers but only the hashes of them. Wickr has received four and a half stars on the Apple store from 80 different ratings and has a similar rating on the Google Play store. The C.E.O of Wickr, Nico Sell has some big plans for Wickr and it will be interesting to see how this company develops in the following years.

BlackBerry is a Canadian based company specializing on mobile communications. It once was the standard smart phone in Canada but quickly consumers switched to over to Apple and Android. Their were rumours about BlackBerry being possible bought by various companies. The company hired a new C.E.O. who changed the focus away from the consumer market. BlackBerry now focuses on one of its strengths which is security and even Preseident Obama uses one.(Marks, 2014, para. 3-4) The BlackBerry Balance which is a software on their phones that allows for the seperation of personal data and work-related data. BlackBerry Balance provides security of the user's information while maintaing a user-friendly design. BlackBerry is a good example of a company focusing on the under supplied market of security and being rewarded by it. The stock price seems to be trending upwards. BlackBerry's stock price is still no where in comparison to their best years of 2008-2009 but maybe their focus on security can help them compete with larger smartphone companies.

\section{Conclusion}\label{sec:conclusion}
There are rising concerns of consumers about privacy issues. Security is often over-shadowed by usability and convenient apps. Legislation is starting to approached these issues related to digital privacy and setting expectations on business. The business world is too focused on big data and profitting from it. This has resulted in companies collecting a large amount of data which it relatively insecure. The increase in security attacks has resulted in a more discussion of levels of security. There are companies starting to market their products as secure devices. There are numerous available technologies available to enable developers to create systems ensure privacy. Developers need to recognize that consumers demand digital privacy and design systems that support anonymity and pseudonymity.

@@

\bibitem{Cook}
	Cook, J. (2014, November 27). The European Parliament Just Voted To Break Up Google Read more: Http://www.businessinsider.com/european-parliament-voted-to-break-up-google-2014-11ixzz3KhLx0CEU. Retrieved November 28, 2014, from http://www.businessinsider.com/european-parliament-voted-to-break-up-google-2014-11

@@

>>>>

\begin{abstract}
Do people care about digital privacy and how can cryptography keep digital information private.
\end{abstract}

@@

The development of databases, the growth of the mobile industry  and the increase usage of the Internet have resulted in a huge collection of personal data. More companies are moving their Information Techonology to the cloud and are creating Data Centers which store an inconceivable amounts of data. Facebook has claimed that they have over a 100 petabytes of photos and video (Wallbank, 2012, para. 8). This along with a strong developer community has lead to anyone being able to create a web application. There are tons of web based hosting companies such as Heroku which allows any to publish their web application easily. New technologies are created yearly for all kinds of web development. There are over 250,000 web related repositories on Github alone. There are numerous plugins that enable your web application to be integrated with numerous social media websites such as Twitter and Facebook. The Internet has provided connections across the world but these connections have brought along some issues.

@@

People are constantly giving out personal information such as email, address, and credit card information daily through various websites. All kinds of information is being tracked when browsing the Internet through Google analytics. With all this information being shared and stored everyday, are people concerned with its security and their privacy?

@@

Business are driven by profit and will look for ways to maximize profit in any manner, if there is not government intervention. Sadly, there is a lack of legislation dealing with the Internet. This is largely due to the how fast web techonologies have grown and how slow the political systems moves. Most of contract law is obsolete when it comes to the Internet and Internet transactions. One example is Canada's Anti-Spam Legislation (CASL) which was passed in Decemeber 2010 but not enforced until July 1, 2014(Government of Canada, 2014, para. 1). It took about a decade and a half to create a spam law and another four years to implement it. The lack of legislation allows for companies to freely deal with their data, and user agreements with little government interference. The goverment is doing little to ensure digital privacy and in certain cases is exploiting personal information in the name of security. A recent example is the Federal Buraeu of Investigation's complaint about Apple's Iphone encryption of data where the user's password was used to encrypt the data (Schneier, 2014, para. 3-4). Encryption is more of a hassle and back doors make their job much easier.

All business have the incentive to improve profit and without external incentives will strictly focus on profits. The two greatest external forces on corporations are the government and consumer purchasing power. The government's pressure on companies is non-existence and what force that is present is outdated. The public is less concerned about with drawbacks of the data privacy and more concerned with the benefits of the application. Most people ask what can it do for me instead of how does it work and what security measures are in place? The lack of incentives have lead to businesses tending to focus more on what profit can be produced from personal data and not the long term impact that it could have. There is a lack of research done by the companies producing these techonologies and their possible consequences on their users. The negative side effects are not taken into consideration during the design process. Security is usually an after thought and privacy is never considered. This is evident in a recent study done by Cisco Canada in which they found ``40 per cent of about 500 firms surveyed had security strategies'' (Blackwell, 2014, para. 2). Just last week, Sony was hacked where a significant amount of employee information was leaked. Their are estimates of ``100 terabytes of data'' were leaked containing information such as ``list of employee salaries and bonuses; Social Security numbers and birth dates; HR employee performance reviews, criminal background checks and termination records'' (Zetter, 2014, para 15-16).

@@

The culture change started by the Internet has brought new social issues. Now a citizen is evaluated by their on-line social interaction. Companies will use Facebook, LinkedIn and Google to screen out potential candidates. Employers need to monitor the social lives closely because of this. One cyber attack can greatly damage one's reputation and may threaten their employment or even their employability. The other thing is the Internet does not forget about one's past. A good example of this is the recent firing of a Quebec teacher: ``the availability on the Internet of erotic films in which she acted created an entirely new context that was not ideal for our students'' (Peritz, 2014, para. 8) was the reason for her dismissal. A teacher's work that occurred over forty years ago should not be revelant to the evaluation of their teaching. This would not have been an issue if the students had just heard about it since it would just be a rumour. The fact that the video is available on the Internet is what the school claims to be the issue. The teacher has no control over the distribution of the content now. This is punishing someone for something that happen a long time ago in which they have no control over it. This problem will only get larger as more information is stored and moved into an age where all data becomes digitalized. It is a frightening fact that all mistakes in one's lifetime will not be forgotten and could impact someone's reputation world wide even forty years later.

The new buzz word in the business world is big data and processing big data for applications. The applications are used by business to make decision and to help better market their product. There is a big push for all companies to collect data and analyzing this data to get ahead. The collection of personal data is usually supported by the possibilities that big data analytics could provide. Their are large claims by people of all the potential benefits including big data makes people happy (Banayan, 2014, para. 4). The collection of this data has little regulations and with an increasing financial incentive it can or may cause some business to collect this information in unethical ways. The other problem is that there is a over confidence in this new field. Micheal Jordan a professor at the University of California stated that ``When you have large amounts of data, your appetite for hypotheses tends to get even larger. And if it’s growing faster than the statistical strength of the data, then many of your inferences are likely to be false. They are likely to be white noise.'' (Gomes, 2014, para. 30) . Google Flu Trends is a system Google developed for tracking the flu. ``Google Flu Trends is often held up as an exemplary use of big data''. (Lazer, 2014, pg. 1). It was reported that Google Flu Trend's prediction were twice as much as the proportion of actual doctor vists for influenze to the Center for Disease Control and Prevention (Lazer, 2014, pg. 1). This is a good case example of how big data does not have all the answers and that it is a field that needs improvement and reasearch. Over confidence in it could lead to costly misled decisions based on ``big data hubris'' (Lazer, 2014, pg.1).

@@

\section{Ways to  Change}\label{sec:developers}
It is clear that security and digital privacy need to have more focus and be strengthened. A developer needs to recognize this during the design of their system. They need to build into their system ways that users can stay annonymous or delete private information when requested. Digital privacy should be the standard when creating web applications and not a feature.

@@

The steps of the system starts by the user creating a note number called $n$ which is a random number. The users then creates a second random number $r$. The note number $n$ is then blinded by combining $n$ with $r$. This is done mathematically by $n * r^b$ where $b$ is the bank's public power which equates to one dollar. The blinded note is then transferred to the bank. The banks signs the blinded note by using its private power $\overline{b}$. The result is returned and the signed note number is obtained by dividing out the $r$ to obtain $n^{\overline{b}}$. The note can be verified that is a signed special number (Chaum, 1985, pg. 7-8).

Banks would need to keep track of transactions to ensure that no note has been used twice. This is done by having a list of all accepted transactions. A transaction is only allowed if it is not part of the list and has the appropriate signature. The underlying digital signature techniques make it difficult for anyone to forge the bank's signature $\overline{b}$. If $r$ is not disclosed then the transcation deposit cannont be linked to the withdrawel (Chaum, 1985, pg. 8-9).

@@

These anonymous remailer aim to allow users to communicate through e-mail without losing privacy. According to Ian Goldberg, David Wagner, and Eric Brewer there are four different versions of Type X Remailers, Type O, Type I, Type II and Type III. Each type provides addtional functionality than the previous version. All the different versions work differently where some are anonymity systems and others are pseudonym systems.

@@

The basic idea of Onion Routing is to create an indirect path between nodes starting at the user's node and ending at the recipents. Each node in the path does not know the recipent but rather just the next node in the path. Each node does not know the path due to the way the packet is encrypted. Each node removes a level of encryption and determines which node to pass to. This path may be used in reverse so the recipent can respond which allows for pseudonymous communications. (Reed, Syverson, \& Goldschag, 1998, pg. 2)

@@

The protocol first exchanges keys using a key exchange. The data is then encrypted using a specified encyrption function such as AES. A message authencating code (MAC) is also created. The encyrpted text and the MAC is sent. The encyrption does not allow for eavesdroppers to read the intended message. The MAC does not allow the message to be intercepted and changed (Sidhpurwala, 2013, para.3).

@@

 Transport Layer Security is widely used. Transport Layer Security provides an example of when standards are created then people will follow them. It is not difficult to enable a web application to support Transport Layer Security through various libraries such as OpenSSL. However, it also is an example of how security budgets are inadequate when compared to their level of importance. This was seen in the HeartBleed bug which occurred in the OpenSSL library this year. OpenSSl is used widely but developed and maintained by a small community of people. It has been stated that ``66 percent of all web servers'' run OpenSSL but OpenSSL Software Foundation receives only ``\$2,000 a year in donations'' (Warren, 2014, para. 28-30). If the support of OpenSSL was comparable to the number of users of the software then the bug would have been much less likely to occur.

\section{Moving Forward}\label{sec:moving-forward}
There are many more privacy enhancing technologies besides the four discussed previously. There is a vast collection of libraries that developers can leverage to help them make their applications more secure. In the early phases of design developers need to focus on security and how their applications deal with privacy issues. A developer has a responsibility to their users to ensure their information is secure and that the user has control over thier private information.

@@

Businesses need to create strategies to handle security and make clear in their user agreements how they store and use information collected. Facebook recently moved towards this by shortening their user agreement and making it more readable for users (Albergotti, 2014, para. 4). This was a step in the write direction but still more needs to be done. Business needs to increases security budgets which are too small. Some companies are seeing this need, which was evident in the IAPP study that reported ``38 percent of respondents said they would likely increase their privacy budget in the next year'' (Solove, 2014, para. 2). There are some companies who have focused on creating secure systems.

\section{Businesses focusing on Security}\label{sec:real-life}
Wickr is a new messenger which is available on Google Play store, Apple App Store, and even for the desktop. It is uses peer-to-peer encryption and enables users to send encrypted messages.  According to its website it uses hashed device information with SHA256 and encrypts all data with AES256. The Wickr server only handles encrypted messages and delete the message upon delivery. The server also never directly interacts with devices identifiers but only the hashes of them (Wickr, 2014, para.1). Wickr has received four and a half stars on the Apple store from 80 different ratings and has a similar rating on the Google Play store. The C.E.O of Wickr, Nico Sell has some big plans for Wickr and it will be interesting to see how this company develops in the following years.

BlackBerry is a Canadian based company specializing on mobile communications. It once was the standard smart phone in Canada but quickly consumers switched to over to Apple and Android. Their were rumours about BlackBerry being possibly bought by various companies. The company hired a new C.E.O. who changed the focus away from the consumer market. BlackBerry now focuses on one of its strengths which is security and even Preseident Obama uses one.(Marks, 2014, para. 3-4) The BlackBerry Balance which is a software on their phones that allows for the seperation of personal data and work-related data. BlackBerry Balance provides security of the user's information while maintaing a user-friendly design. BlackBerry is a good example of a company focusing on the under supplied market of security and being rewarded by it. The stock price seems to be trending upwards. BlackBerry's stock price is still no where in comparison to their best years of 2008-2009 but maybe their focus on security can help them compete with larger smartphone companies.

\section{Conclusion}\label{sec:conclusion}
There are rising concerns of consumers about privacy issues. Security is often over-shadowed by usability and convenient apps. Legislation is starting to approached these issues related to digital privacy and setting expectations on business. The business world is too focused on big data and profitting from it. This has resulted in companies collecting a large amount of data which is relatively insecure. The increase in security attacks has resulted in a more discussion about the levels of security. There are companies starting to market their products as secure devices. There are numerous available technologies available to enable developers to create systems ensure privacy. Developers need to recognize that consumers demand digital privacy and design systems that support anonymity and pseudonymity.

@@

\bibitem{Cook}
	Cook, J. (2014, November 27). The European Parliament Just Voted To Break Up Google. Retrieved November 28, 2014, from http://www.businessinsider.com/european-parliament-voted-to-break-up-google-2014-11

@@

\bibitem{Wickr}
	Wickr, (2014). How Wickr Works. Retrieced Devember 7, 2014, from https://www.wickr.com/how-wickr-works/
